\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2013-12-28 doc/paper.tex
extend EWSB solving section

\section{Stucture of the spectrum generator}
\subsection{boundary conditions}
Different types, how we match to low energy data.  reference EWSB or merge that section here.
mention two loop qcd parts from softsusy.

@@

\subsection{tree level spectrum}
The diagonalised tree level $\overline{DR}$ masses are obtained from expressions...
The numerical diagonalisation is performed by... with an accuracy set by...   
\subsection{Electroweak symmetry breaking}
At tree level the EWSB conditions are written...
The one loop electroweak symmetry breaking are solved using the gsl routines finding multiple roots, \code{gsl_multiroot_fsolver}.  Iteration. For example in the MSSM \fs generates.. show code.
 If higher accuracy is required additional routines with higher order corrections can be added in a very simple way.  For example in the MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of $\oatas$, $\oabas$, $\oatq$, $\oabatau$, 
$\oabq$, $\oatauq$ and $\oatab$  as follows...  
Something similar is done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}.
 Corrections from other sources can be added to the generated code by...
\subsection{two scale fixed point iteration}
The default solver which calls the above routines above and finds the full set of $\overline{DR}$ masses in the model consistent with all constraints is the two-scale solver.  This can be selected by... but if no selection is made this will be used anyway as default.

>>>>

\section{Stucture of the spectrum generator}

@@

\subsection{Boundary conditions}
In \fs the RGE boundary conditions are implemented by inheriting from
a common \code{Constraint<Two\_scale>} interface class and
implementing the pure virtual functions \code{apply()},
\code{get\_scale()} and \code{set\_model()}, see
\figurename~\ref{fig:schematic-two-scale-constraint-interface}.  Per
default, \fs creates three default boundary conditions:
%
\begin{itemize}
\item \emph{GUT constraint:} Here the \code{apply()} function
  calculates the value of the unification scale $M_X$ (from the
  condition given in \code{HighScale}) and sets the model parameters
  given in \code{HighScaleInput}.
\item \emph{SUSY constraint:} Here the \code{apply()} function
  calculates the value of the susy scale $M_S$ and sets the model
  parameters given in \code{SUSYScaleInput}.  Afterwards the EWSB
  conditions are solved for the parameters given in
  \code{EWSBOutputParameters} at the one-loop level, i.e., the the
  parameters in \code{EWSBOutputParameters} are determined such that
  the one-loop effective potential is minimized.
\item \emph{Low-scale constraint:} This constraint matches the gauge
  and Yukawa couplings of the susy model to the Standard Model values
  using threshold corrections.  Furthermore the parameters given in
  \code{LowScaleInput} are set.
\end{itemize}

{\color{red}Different types, how we match to low energy data.
  reference EWSB or merge that section here.  mention two loop qcd
  parts from softsusy.}

\subsubsection{Electroweak symmetry breaking}
In \fs the one-loop electroweak symmetry breaking conditions are
formulated as
%
\begin{align}
  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEVs corresponding to the Higgs field $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.

The electroweak symmetry breaking conditions
\eqref{eq:one-loop-ewsb-eq} are then solved simultaneously using the
iterative multi-dimensional root finder algorithm
\code{gsl_multiroot_fsolver_hybrid} from the Gnu Scientific Libraray
(GSL).  If no root was found, the \code{gsl_multiroot_fsolver_hybrids}
algorithm is tried as alternative, which uses a variable step size but
might be a little slower.

For example in the MSSM \fs expresses \eqref{eq:one-loop-ewsb-eq} as
the function
%
\begin{lstlisting}[language=C++]
int MSSM<Two_scale>::tadpole_equations(const gsl_vector* x, void* params, gsl_vector* f)
{
   ...

   const CLASSNAME::Ewsb_parameters* ewsb_parameters
      = static_cast<CLASSNAME::Ewsb_parameters*>(params);
   MSSM* model = ewsb_parameters->model;
   const unsigned ewsb_loop_order = ewsb_parameters->ewsb_loop_order;

   double tadpole[numberOfEWSBEquations];

   model->set_BMu(gsl_vector_get(x, 0));
   model->set_Mu(INPUT(SignMu) * Abs(gsl_vector_get(x, 1)));

   // calculate tree-level tadpole eqs.
   tadpole[0] = model->get_ewsb_eq_vd();
   tadpole[1] = model->get_ewsb_eq_vu();

   // subtract one-loop tadpoles
   if (ewsb_loop_order > 0) {
      model->calculate_DRbar_parameters();
      tadpole[0] -= Re(model->tadpole_hh(0));
      tadpole[1] -= Re(model->tadpole_hh(1));

   }

   for (std::size_t i = 0; i < numberOfEWSBEquations; ++i)
      gsl_vector_set(f, i, tadpole[i]);

   return GSL_SUCCESS;
}
\end{lstlisting}
%
Here \code{x} is the vector of EWSB output parameters
(\code{EWSBOutputParameters}) and \code{f} is a vector which contains
the one-loop EWSB eqs.\ \eqref{eq:one-loop-ewsb-eq}.  This
\code{tadpole_equations()} function is called inside of
\code{MSSM<Two_scale>::solve_ewsb_iteratively_with} then via
%
\begin{lstlisting}[language=C++]
// initial guess
double x_init[numberOfEWSBEquations];
ewsb_initial_guess(x_init);

// setup root finder
int ewsb_loop_order = 1;
Ewsb_parameters params = {this, ewsb_loop_order};
Root_finder<numberOfEWSBEquations> root_finder(MSSM<Two_scale>::tadpole_equations,
                           &params,
                           number_of_ewsb_iterations,
                           ewsb_iteration_precision);
root_finder.set_solver_type(gsl_multiroot_fsolver_hybrid);

root_finder.find_root(x_init);
\end{lstlisting}

If higher accuracy is required additional routines with higher order
corrections can be added in a very simple way.  For example in the
MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by
P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of
$\oatas$, $\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and
$\oatab$ as follows...  Something similar is done for the NMSSM, but
in this case the NMSSM $\oatas$, $\oabas$ pieces come from
\cite{Degrassi:2009yq}.  Corrections from other sources can be added
to the generated code by...

\subsection{tree level spectrum}
The diagonalised tree level $\overline{DR}$ masses are obtained from expressions...
The numerical diagonalisation is performed by... with an accuracy set by...   
\subsection{two scale fixed point iteration}
The default solver which calls the above routines above and finds the full set of $\overline{DR}$ masses in the model consistent with all constraints is the two-scale solver.  This can be selected by... but if no selection is made this will be used anyway as default.
